---
title: "HW 1"
author: "Wayne Monical"
date: "2024-09-20"
output: github_document
---

Here we load the `penguins` data set from the `palmerpenguins` package.

```{r}
data("penguins", package = "palmerpenguins")
```


This data set comes in the form of a data frame. It contains data on three different penguin `species`, Adelie, Gentoo, and Chinstrap. The penguins range over three different islands, Torgersen, Biscoe, and Dream, given by the `island` variable. Observations come over three years, 2007 through 2009, given by the `year` variable.

The data describes `r nrow(penguins)` penguin individuals in a total of `r ncol(penguins)` variables. 
The data includes various physical measures of the penguins, such as their flipper length in millimeters, given by the `flipper_length_mm` variable. This variable has an average value of `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` in this data set. 

```{r}
print(unique(penguins$species))
print(unique(penguins$island))
print(unique(penguins$year))
```


```{r}
nrow(penguins)
```


```{r}
head(penguins)
```

